---
title: "NYC Flight Delay Report"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nycflights13)
library(tidyverse)
library(janitor)
flights <- nycflights13::flights %>%
    remove_constant() %>%
    drop_na(dep_time)

```

## Introduction

We will use the library nycflights13 which contains information about 16 different airline flying from 3 different airports in New York City.
It contains 5 datasets that provide info about flights, airports, weather, planes, airlines.
This data set was presented by the bureau of transportation and statistics.
Through more research we’ve found out that there are actually more than 125 airlines but we couldn’t find out why they choose these 16.

## Exploratory Data Analysis

### Glimpsing the data

We used the glimpse function to get a quick look at the data.

glimpse(flights)

We can see that the flights dataset has 328,521 rows and 19 columns.
We removed the years column because it has the same value for all the rows, and we removed the rows that had missing values for the departure time.

## Exploratory Data Analysis
### Check which airport has the most flights
```{r}
flights %>%
    count(origin) %>%
    arrange(desc(n))
```

### Check which airport has the most delays
```{r}
flights %>%
    filter(dep_delay > 0) %>%
    count(origin) %>%
    arrange(desc(n))
```

## Exploratory Data Analysis
### Check the proportion of delay to the number flights

```{r}
flights %>%
    filter(dep_delay > 0) %>%
    count(origin) %>%
    mutate(prop = n / sum(n)) %>%
    arrange(desc(prop))
```

## Key investigations results from the data
### Which month has the most delays?

```{r}
flight %>%
  filter(dep_delay > 0) %>%
  group_by(month, origin) %>%
  summarise(count=n()) %>%
  ggplot(aes(month, count, color = origin))+
  geom_point(shape = 4) +
  geom_line() +
  scale_x_continuous("Month", n.breaks = 12) +
  scale_y_continuous("Count of delays") +
  ggtitle("Delays count by month from all NYC airport.") +
  theme_classic()
```

## Key investigations results from the data
### What is the distribution of all delays?

```{r}
ggplot(flight, aes(x = dep_delay)) +
  geom_histogram(bins = 80) +
  scale_x_continuous(limits = c(-100, 500))
```

## Key investigations results from the data
### What is the average delay for each airline?

```{r}
flight %>%
  group_by(carrier) %>%
  summarise(avg = mean(dep_delay), std = sd(dep_delay)) %>%
  ggplot(aes(carrier, avg)) +
  geom_col() +
  #geom_errorbar(aes(ymin=avg-std, ymax=avg+std))+
  scale_y_continuous("Average delay in min") +
  scale_x_discrete("Carrier") +
  ggtitle("Average delay in minute for each carrier thorughout 2013") +
  theme_classic()
```

## Takeaways

* Most flights and delays happen during the summer months
* The proportion of delays increases with the number of flights
* F9 has the highest average delay

